
%
% intro
%

\chapter{Introduction}


\section{Motivation}

Digital mapping applications on the Internet are strongly emerging. Big players like Google Maps\footnote{\url{https://maps.google.at}} and OpenStreetMap\footnote{\url{http://www.openstreetmap.org/ }} provide online maps, that users can view and interact with. 

Maps allow telling stories and communicating data in a visual way. Using open source tools such as TileMill\footnote{\url{http://mapbox.com/tilemill}} and online services like CloudMade\footnote{\url{http://cloudmade.com/ }}, more and more people are able to create their own custom maps. The Free and Open Source content management system and framework Drupal\footnote{\url{http://drupal.org}} provides tools for adding, editing and visualizing geographic data on maps. This allows to integrate interactive map applications into web sites.

Maps provide a birds-eye-view on the geography and information captured in such system. Sometimes, they are used to get a quick overview of points-of-interest in a certain area. If a large amount of information is contained in such an area, problems in terms of computability and visual clutter arise: visualizing thousands of points on a single map both challenge human and computers. Obviously, when telling a story, information needs to be told in a compact way as the human brain can only process a limited amount of data at the same time. Similarly, large amounts of data involve a higher burden on the computer components that participate in the web mapping process~~\cite{noellenburg11geovis, Delort10vis}.

Clustering\footnote{\url{http://en.wikipedia.org/wiki/Cluster_analysis}} is a technique for grouping objects with similarities that can be used to reduce visual clutter as described before. Various client-side Javascript libraries like Leaflet.markercluster\footnote{\url{https://github.com/Leaflet/Leaflet.markercluster}} exist for clustering points on maps. This enhances performance and readability of data-heavy map applications. But still, all data needs to be transferred to the client and processed on a potentially slower end user device. By clustering data on the server-side, the load is shifted from the client to the server which allows displaying larger amounts of data in a performant way. Professional services like maptimize\footnote{\url{http://www.maptimize.com}} provide such a functionality, while in the open source space little libraries and frameworks exist for server-side clustering of geospatial data.

In order to create interactive maps based on large data sets, this thesis evaluates and implements a performance-optimized server-side clustering algorithm for Drupal.


\section{Outline of the thesis}

Chapter \ref{chapter:foundations} introduces the technical foundations for this thesis. It provides an overview by explaining cluster theory and comparing four main clustering algorithms. Further, a discussion on spatial data computation introduces space order and space decomposition methods, Quadtrees and the Geohash encoding. Finally, an overview of foundational concepts in web mapping as coordinate systems, map projections and spatial data types is provided. 

Chapter \ref{chapter:state} discusses the state of the art of related technologies for the thesis. An explanation of a modern web mapping stack is given as well as the basics of Drupal \& mapping technologies. Further, the state of the art for client-side and server-side clustering technologies in web mapping is analyzed.

Chapter \ref{chapter:objectives} states the objectives and exemplary use cases for the thesis. It defines the goals to be accomplished as a result of the implementation part of the thesis.

Chapter \ref{chapter:realization} describes the realization of the server-side clustering algorithm. First, an analysis based on the objectives stated in the previous chapter is given. Subsequently, the Geohash-based clustering algorithm is defined. Finally, the architecture and implementation of the algorithm for Drupal is explained in detail.

Chapter \ref{chapter:use-cases} discusses the implementation of use cases for the realized server-side clustering algorithm.

Chapter \ref{chapter:conclusions-outlook} finally evaluates the results of the thesis. It contains performance tests and a further evaluation on how the defined objectives have been accomplished. Final conclusions are made and an outlook on future work is given.




